% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  ignorenonframetext,
]{beamer}
\title{Introduction to Data Science}
\author{Stat 220 Bastola}
\date{2022-01-06}

\usepackage{pgfpages}
\setbeamertemplate{caption}[numbered]
\setbeamertemplate{caption label separator}{: }
\setbeamercolor{caption name}{fg=normal text.fg}
\beamertemplatenavigationsymbolsempty
% Prevent slide breaks in the middle of a paragraph
\widowpenalties 1 10000
\raggedbottom
\setbeamertemplate{part page}{
  \centering
  \begin{beamercolorbox}[sep=16pt,center]{part title}
    \usebeamerfont{part title}\insertpart\par
  \end{beamercolorbox}
}
\setbeamertemplate{section page}{
  \centering
  \begin{beamercolorbox}[sep=12pt,center]{part title}
    \usebeamerfont{section title}\insertsection\par
  \end{beamercolorbox}
}
\setbeamertemplate{subsection page}{
  \centering
  \begin{beamercolorbox}[sep=8pt,center]{part title}
    \usebeamerfont{subsection title}\insertsubsection\par
  \end{beamercolorbox}
}
\AtBeginPart{
  \frame{\partpage}
}
\AtBeginSection{
  \ifbibliography
  \else
    \frame{\sectionpage}
  \fi
}
\AtBeginSubsection{
  \frame{\subsectionpage}
}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usetheme[]{metropolis}
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Introduction to Data Science},
  pdfauthor={Stat 220 Bastola},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\newif\ifbibliography
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\def\begincols{\begin{columns}}
\def\begincol{\begin{column}}
\def\endcol{\end{column}}
\def\endcols{\end{columns}}

\setbeamercolor{structure}{fg=blue!55!green}
\usepackage{graphicx}
\usepackage{mathabx}
\usepackage{xcolor}


%% change fontsize of R code
\let\oldShaded\Shaded
\let\endoldShaded\endShaded
\renewenvironment{Shaded}{\tiny\oldShaded}{\endoldShaded}

%% change fontsize of output
\let\oldverbatim\verbatim
\let\endoldverbatim\endverbatim
\renewenvironment{verbatim}{\tiny\oldverbatim}{\endoldverbatim}

%\usepackage[sfdefault]{FiraSans}

% declare overall beamer theme to use as baseline
\usetheme{default}


% make code-output smaller
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{fontsize=\tiny,commandchars=\\\{\}}

% make console-output smaller:
  \makeatletter
\def\verbatim{\tiny\@verbatim \frenchspacing\@vobeyspaces \@xverbatim}
\makeatother


\setlength{\parskip}{1pt}


\setlength{\OuterFrameSep}{0pt}
\makeatletter
\preto{\@verbatim}{\topsep=1pt \partopsep=1pt }
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\begin{document}
\frame{\titlepage}

\begin{frame}{Something about me}
\protect\hypertarget{something-about-me}{}
\begincols
  \begincol{.60\textwidth}
\begin{itemize}
    \item First year at Carleton
    \item Originally from Nepal
    \item PhD in Applied Statistics from UC-Riverside
    \item Diverse education background
    \item Avid learner and traveler
\end{itemize}

\endcol
\begincol{.40\textwidth}

\endcol \endcols
\end{frame}

\begin{frame}{COVID-19 related policies}
\protect\hypertarget{covid-19-related-policies}{}
\begin{itemize}
\item Stay home when sick. (Even if you don’t have COVID-19, you should stay home if you aren’t feeling well.)
\item Follow \href{https://www.cdc.gov/coronavirus/2019-ncov/vaccines/fully-vaccinated.html}{CDC} on testing, quarantine, and isolation.
\item Follow the College mask-wearing policy
\end{itemize}
\end{frame}

\begin{frame}{What is data science?}
\protect\hypertarget{what-is-data-science}{}
\textbf{Data Science}:

\begin{itemize}
\tightlist
\item
  the science of extracting meaningful information from data
\end{itemize}

\vspace*{0.7in}

\begin{quote}
Computer science is more than just programming; it is the creation of
appropriate abstractions to express computational structures and the
development of algorithms that operate on those abstractions. Similarly,
statistics is more than just collections of estimators and tests; it is
the interplay of general notions of sampling, models, distributions, and
decision-making. {[}Data science{]} is based on the idea that these
styles of thinking support each other. - Michael Jordan, UC Berkeley
\end{quote}
\end{frame}

\begin{frame}{Data Scientist}
\protect\hypertarget{data-scientist}{}
The ``data scientist'' mashup:

\begin{itemize}
\tightlist
\item
  ``The definitions of data science are converging around the
  intersection of mathematics, statistics, and computer science---with
  some area of application (e.g., finance, biology, political
  science).''
\item
  ``I have heard data scientists referred to equally as

  \begin{itemize}
  \tightlist
  \item
    `the computer scientist who was the best of his peers in his
    statistics courses' and
  \item
    `the statistician who was the best of his peers in his computer
    science courses.'''
  \end{itemize}
\end{itemize}

\begin{quote}
-Jennifer Lewis Priestley
\href{http://magazine.amstat.org/blog/2016/01/01/data-science-the-evolution-or-the-extinction-of-statistics/}{Data
Science: The Evolution or the Extinction of Statistics?}
\end{quote}
\end{frame}

\begin{frame}{Data Science Education}
\protect\hypertarget{data-science-education}{}
\begin{itemize}
\tightlist
\item
  Many schools now offer
  \href{http://datascience.community/colleges}{degrees} in some form of
  data science (data analytics)
\item
  A B.S. (or Masters) in Data Science includes courses like:

  \begin{itemize}
  \tightlist
  \item
    Intro Stats, Intro Programming, Intro Data Science
  \item
    Regression (modeling)
  \item
    Machine Learning, Data mining
  \item
    Database management
  \item
    Data visualization
  \item
    Big Data
  \item
    Applications (econ, poli sci, bio)
  \item
    Ethics
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Stat 220: Data Science}
\protect\hypertarget{stat-220-data-science}{}
Focus on the ``soup to nuts'' approach to problem solving

\begin{itemize}
\tightlist
\item
  data wrangling

  \begin{itemize}
  \tightlist
  \item
    reshaping, cleaning, gathering
  \end{itemize}
\item
  learning from data

  \begin{itemize}
  \tightlist
  \item
    EDA tools
  \item
    statistical learning methods
  \item
    network data, spatial data
  \end{itemize}
\item
  communication

  \begin{itemize}
  \tightlist
  \item
    reproducibility
  \item
    effective visualization
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{How to make friends and succeed in Data Science?}
\protect\hypertarget{how-to-make-friends-and-succeed-in-data-science}{}
\begin{enumerate}
\tightlist
\item
  \emph{Actively follow along!} RMarkdown documents will be provided for
  you each week - use these to take notes and run code ``live'' in
  class.
\item
  \emph{Ask questions!} This is new for everyone, no question is a bad
  question.
\item
  \emph{When you don't know if something will work, try it!}
  Experimentation is key in this class.
\item
  \emph{BRING YOUR LAPTOP.} Charged, everyday.
\end{enumerate}
\end{frame}

\begin{frame}{Tell me something about yourself}
\protect\hypertarget{tell-me-something-about-yourself}{}
\begincols
  \begincol{.60\textwidth}
\begin{itemize}
\item Your name?
\item Gender Pronouns?
\item Why are you interested in data science?
\item Rate your R Skills from 1 to 10.
\end{itemize}
\endcol
\begincol{.40\textwidth}

\endcol \endcols
\end{frame}

\begin{frame}[fragile]{Class Instruction}
\protect\hypertarget{class-instruction}{}
\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Main Course Webpage for Instruction}
\end{Highlighting}
\end{Shaded}

\center{\url{https://deepbas.io/courses/stat220/}}

\begin{itemize}
\tightlist
\item
  Bookmark this page
\item
  Should be checked multiple times a day
\end{itemize}
\end{frame}

\begin{frame}{Collaborative notes}
\protect\hypertarget{collaborative-notes}{}
\begin{itemize}
\item Each day, two of you will collaborate on notes to share with the class
\item Creates a crowd-sourced version of what we do in class
\item Helps anyone who needs to miss class
\item You'll do this 3x throughout the course
\item \href{https://drive.google.com/drive/folders/1xgSlButXtTHNvZCi9QkhOv3YSBDeB1UH?usp=sharing}{Sign up here}
\item Notes are due 24 hours after class, count as a HW assignment
\end{itemize}
\end{frame}

\begin{frame}{Necessary skills to be mastered}
\protect\hypertarget{necessary-skills-to-be-mastered}{}
\begin{itemize}
\item
  programming with data
\item
  statistical modeling
\item
  domain knowledge
\item
  communication
\end{itemize}
\end{frame}

\begin{frame}{Why aren't we learning Python?}
\protect\hypertarget{why-arent-we-learning-python}{}
When Hadley Wickham was asked ``Why R?''

\vspace*{0.5in}

\begin{quote}
And the second reason, which is both a huge strength of R and a bit of a
weakness, is that R is \textbf{not just a programming language}. It was
designed from day 1 to be an \textbf{environment that can do data
analysis}. So, compared to the other options like Python, you can get up
and running in R doing data science, learning much, much less about
programming to get started. And that generally makes it like
\textbf{easier to get up and running if you don't have formal training
in computer science or software engineering}.
\end{quote}

\begin{quote}
-Hadley Wickham,
\href{https://www.r-bloggers.com/advice-to-young-and-old-programmers-a-conversation-with-hadley-wickham/}{Advice
to Young (and Old) Programmers: A Conversation with Hadley Wickham}
\end{quote}
\end{frame}

\begin{frame}[fragile]{Using R Markdown for data science}
\protect\hypertarget{using-r-markdown-for-data-science}{}
\begin{itemize}
\tightlist
\item
  You will use \href{http://rmarkdown.rstudio.com/lesson-1.html}{R
  Markdown} for all work in this class
\item
  A Markdown (\texttt{.Rmd}) file contains

  \begin{itemize}
  \tightlist
  \item
    R code
  \item
    written answers, description of results, report, etc.
  \end{itemize}
\item
  The Markdown file is \texttt{knit} to generate an output document

  \begin{itemize}
  \tightlist
  \item
    pdf, html, word
  \item
    presentations (html, beamer pdf)
  \item
    dashboards, interactive graphics (html)
  \end{itemize}
\item
  Markdown is designed for \textbf{reproducibility}!
\item
  The slides I produce for this class are R Markdown's
  \href{http://rmarkdown.rstudio.com/beamer_presentation_format.html}{\texttt{beamer}}
\end{itemize}
\end{frame}

\end{document}
