---
title: "MLR Inference"
subtitle: "<br/> STAT 230"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  xaringan::moon_reader:
    css: ["default", css/xaringan-themer-solns.css, css/my-theme.css, css/my-font.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    seal: false
    nature:
      highlightStyle: googlecode  #http://arm.rbind.io/slides/xaringan.html#77 # idea, magula
      highlightLines: true
      highlightLanguage: ["r", "css", "yaml"]
      countIncrementalSlides: true
      slideNumberFormat: "%current%"
      titleSlideClass: ["left", "middle", "inverse"]
      ratio: "16:9"
    includes:
      in_header: header.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(htmltools.preserve.raw = FALSE)
options(ggrepel.max.overlaps = Inf)

knitr::opts_chunk$set(echo = TRUE, 
                      dev = 'svg',
                      collapse = FALSE, 
                      comment = NA,  # PRINTS IN FRONT OF OUTPUT, default is '##' which comments out output
                      prompt = FALSE, # IF TRUE adds a > before each code input
                      warning = FALSE, 
                      message = FALSE,
                      fig.height = 3, 
                      fig.width = 4,
                      out.width = "100%"
                      )

# load necessary packages
library(Sleuth3)   # Data-set for Sleuth
library(tidyverse)
library(dplyr)
library(countdown)
library(mosaic)
library(ggthemes)
library(xaringanExtra)
library(forcats)
xaringanExtra::use_panelset()
xaringanExtra::use_tachyons()
xaringanExtra::use_clipboard()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = TRUE  
)
library(flipbookr)
library(patchwork)
library(DT)
library(moderndive)
library(knitr)
library(grid)
library(gridExtra)
library(Stat2Data)
library(palmerpenguins)

# Set ggplot theme
theme_set(theme_tufte(base_size = 10))

yt <- 0

set.seed(1234)

sim_mlr = function(x1, x2, beta_0 = 2, beta_1 = 3, beta_2 = 2, sigma = 3) {
  n = length(x1)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x1 + beta_2 * x2 + epsilon
  data.frame(predictor1 = x1, predictor2 = x2, response = y)
}

num_obs = 25
x_vals1 = runif(num_obs, 0, 10)
x_vals2 = rbeta(num_obs, 1,5)*10
sim_data = sim_mlr(x1 = x_vals1, x2 = x_vals2, beta_0 = 3, beta_1 = 2, beta_2= 1.6, sigma = 3)


```


```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


layout: true
  
---

class: title-slide, middle

# .fancy[MLR Inference]

### .fancy[Stat 230]

`r format(Sys.Date(), ' %B %d %Y')`

---

# Overview

.pull-left[
```{r, echo=FALSE, fig.align='center', fig.width=4, fig.height=4, out.width="100%"}
library(gg3D)
ggplot(sim_data, aes(predictor1, y=predictor2, z=response)) + 
  theme_void()+
  axes_3D() +
  stat_3D()
```

]
.pull-right[
<br>
<br>
Today: 
.blockquote-list[
MLR Inference
- t tests/CIs for coefficients
- CI for mean response, PI for new response
- Linear combinations of coefficients: CIs and tests
]

]

---

class: middle

# MLR Model

$$Y_{i}=\beta_{0}+\beta_{1} x_{1}+\beta_{2} x_{2}+\cdots \beta_{p} x_{p}+\epsilon_{i}, \epsilon_{i} \sim N(0, \sigma)$$

.blockquote[
MLR inference:
- estimation is the same as SLR!
- tests/CI are also same as SLR!
  - .bold[Except] we use $n-(p+1)$ degrees of freedom where $p+1$ is the number of $\beta$'s in the model (including the intercept)

]


---

class: middle

# MLR inference for one parameter: Testing

.blockquote-list[
- Testing single parameters:
$$\begin{array}{c}
H_{0}: \beta_{j}=0 \quad H_{A}: \beta_{j} \neq 0 \\
t=\frac{\hat{\beta}_{j}-0}{S E\left(\hat{\beta}_{j}\right)} \\
\text{p-value }=2 \times P(T>|t|)
\end{array}$$
using t-distribution with $n-(p+1)$ degrees of freedom
]

---

class: middle

# MLR inference for one parameter: Confidence intervals (CI)

.blockquote[
- CI for a single parameters:
$$\hat{\beta}_{j} \pm t^{*} S E\left(\hat{\beta}_{j}\right)$$
$t^{*}$ is the $(100-C) / 2$ percentile from the t-distribution with $n-(p+1)$ degrees of freedom
]

---

# Example: RECS MLR

.code90[
```{r}
library(moderndive)
energy <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/EnergySurvey.csv")
energy <- energy %>% 
  mutate(logCost = log(CostTotal), 
         logSqft = log(SqftMeasure))

cost_lm_log <- lm(logCost ~ logSqft + HHSize, data = energy)
reg_table_energy <- get_regression_table(cost_lm_log, digits = 4)
knitr::kable(reg_table_energy, format = "html")
```
]

$$\hat{\mu}(\log (\text { Cost }) \mid x)=4.3667+0.3722 \log (\text { Sqft })+0.0839 \text { HHSize }$$


---

# Example: RECS MLR

.code80[
```{r}
get_regression_summaries(cost_lm_log) %>% knitr::kable(digits = 4)
```
]

<br>
--

> What does $H_0: \beta_2 = 0$ vs. $H_A: \beta_2 \neq 0$ test?

---

# Example: RECS MLR

.code80[
```{r, echo=FALSE}
knitr::kable(reg_table_energy[3,], 
              digits = 4, format = "html")
```
]
<br>
.code80[
```{r, echo=FALSE}
get_regression_summaries(cost_lm_log) %>% kable(digits = 4)
```
]

.blockquote.font80[
- Is household size associated with energy cost after controlling for dwelling size?
- Test stat? p-value?
$$t=\frac{0.0839-0}{0.0046} \approx 18.19$$
$$\text {p-value }=2 \times P(T>18.19)=2 \times P(T<-18.19)= 2 \times \operatorname{pt}(-18.19, \mathrm{df}=4381-3)<0.0001$$]

.pull-left[
.code80[
```{r, collapse=TRUE}
# 2 times area in the left tail 
# beyond negative observed test stat
2*pt(-18.19, df = 4378)
```
]
]

.pull-right[
.code80[
```{r, collapse=TRUE}
# 2 times area in the right tail 
# beyond the observed test stat
2*(1 - pt(18.19, df = 4378)) 
```
]
]

---

# Example: RECS MLR

```{r, echo=FALSE}
knitr::kable(reg_table_energy[3,], 
              digits = 4, format = "html")
```

.blockquote.font80[
<br>
Is household size associated with energy cost after controlling for dwelling size?
- Yes, after accounting for dwelling square footage, the estimated effect of household size on mean energy costs is statistically significant $(\mathrm{t}=18.19, \mathrm{df}=4378, \mathrm{p}<0.0001)$.
- Quantify the "effect": $95 \%$ CI for $\beta_{2}$
$$0.0839 \pm(1.9605)(0.0046)=(0.075,0.093)$$
]

```{r, collapse=TRUE}
qt(.975, df = 4378)
0.0839 + c(-1,1)*(1.9605)*(0.0046)
```


---

# Example: RECS MLR

.blockquote.font80[
- Quantify the "effect": $95 \%$ CI for $\beta_{2}$ is $0.075$ to $0.093$
- Our response was logged but HHSize was not (think, SLR exponential model)
]

--

```{r}
cost_lm_log <- lm(logCost ~ logSqft + HHSize, data = energy)
```

--

.blockquote.font80[A 1 person increase in household size is associated with a multiplicative
change in median energy cost between]

```{r, collapse=TRUE}
exp(c(.075, .093))
```

--

.blockquote-list.font80[We are 95% confident that a 1 person increase in household size is associated with 7.9% to 9.7% increase in median energy cost after accounting for dwelling size.]

---

# Example: Pengiuns from `palmerpenguins`

.blockquote-list[
Model mean mass as a function of bill length, species and their interaction:
$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}$$
- Mean function for Adelie (baseline species)
$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2}(0)+\beta_{3}(0)+\beta_{4} b i l l \times(0)+\beta_{5} b i l l \times(0) \\
&=\beta_{0}+\beta_{1} \text { bill }
\end{aligned}$$
- $\beta_{1}$ : effect of bill length on mass for Adelie (baseline)
- .yellow-h[95% CI for this parameter??]
]

---

# Example: Pengiuns from `palmerpenguins`


```{r}
# load library and remove NA cases for the 3 variables of interest
library(palmerpenguins)
library(tidyr) # has `drop_na()` function that drops rows having missing values
penguins <- penguins %>% tidyr::drop_na(bill_length_mm, body_mass_g, species)
peng_interaction_lm <-  lm(body_mass_g ~ bill_length_mm*species, data = penguins)
peng_table_interaction <- get_regression_table(peng_interaction_lm, digits = 5)
knitr::kable(peng_table_interaction, digis= 5, format = "html")
```

---

# Example: Pengiuns from `palmerpenguins`

```{r, echo=FALSE}
knitr::kable(peng_table_interaction[2,], digits= 5, format = "html")
get_regression_summaries(peng_interaction_lm) %>% kable(digits = 5)
```

```{r, collapse=TRUE}
qt(.975, df = 336) # df = 342 - 6 = 336
```


.blockquote[
- A one $\mathrm{mm}$ increase in bill length in Adelie penguins is associated with a $72.1 \mathrm{~g}$ to $116.9 \mathrm{~g}$ increase in mean weight.
$$94.49982 \pm(1.9670)(11.39794)=72.08,116.92$$]


---

# Example: Pengiuns from `palmerpenguins`

.blockquote-list[
Model mean mass as a function of bill length, species and their interaction:
$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}$$
- Mean function for Chinstrap
$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} b i l l+\beta_{2}(1)+\beta_{3}(0)+\beta_{4} b i l l \times(1)+\beta_{5} b i l l \times(0) \\
&=\beta_{0}+\beta_{2}+\left(\beta_{1}+\beta_{4}\right) b i l l
\end{aligned}$$
- $\beta_{1}+\beta_{4}$ : effect of bill length on mass for Chinstrap
- .yellow-h[95% CI for this parameter??]
]

---

class: middle

# Sampling distribution of $\hat{\beta}_j$'s

.blockquote[
Before, we looked at how $\hat{\beta}_{j}$ estimates are correlated 
- For penguins:
- $\hat{\beta}_{1}$ and $\hat{\beta}_{4}$ are also correlated, meaning
$$S E\left(\hat{\beta}_{1}+\hat{\beta}_{4}\right) \neq \sqrt{S E\left(\hat{\beta}_{1}\right)^{2}+S E\left(\hat{\beta}_{4}\right)^{2}}$$
]

---

# New Inference: linear combinations of two $\beta_j$

.bold[Parameter of interest:]

$$\gamma=c_{i} \beta_{i}+c_{j} \beta_{j}$$
where $c_{i}$ and $c_{j}$ are known constants.

.blockquote-list.font90[
- The effect of bill length on mass for Chinstraps:
$$\gamma=(1) \times \beta_{1}+(1) \times \beta_{4}=\beta_{1}+\beta_{4}$$
where $c_{1}=1$ and $c_{4}=1$.
- Holding bill length fixed, $\beta_{2}+\beta_{4} b i l l$ is the difference in mean mass between Chinstrap and Adelie.
- If bill length is $45 \mathrm{~mm}$, this difference is
$$\gamma=(1) \times \beta_{2}+(45) \times \beta_{4}=\beta_{2}+45 \beta_{4}$$
where $c_{2}=1$ and $c_{4}=45$.
]

---

# New Inference: linear combinations of two $\beta_j$

.bold[Estimated parameter:]

$$\gamma=c_{i} \beta_{i}+c_{j} \beta_{j}$$
where $c_{i}$ and $c_{j}$ are known constants.

```{r, collapse=TRUE}
knitr::kable(peng_table_interaction[c(2,5),], digits= 5, format = "html")
```

.blockquote[
- The estimated effect of bill length on mass for Chinstraps:
$$\hat{\gamma}=\hat{\beta}_{1}+\hat{\beta}_{4}=94.49982+(-35.38208)=59.11774$$
]
---

# New Inference: linear combinations of two $\beta_j$

.pull-left-60[
```{r, echo=FALSE}
ggplot(penguins, aes(x = bill_length_mm, 
                     y = body_mass_g, 
                     color = species)) +
theme(legend.position = "bottom")+
geom_point() +
geom_smooth(method = "lm",
             se = FALSE, 
             aes(color=species))+
labs(title = "Regression of mass on bill length and species") 
```
]
.pull-right-40[
.blockquote[
Quick check: does it make sense that the effect of bill length on mass (.bold[slope]) less for Chinstrap than Adelie?
]

]

---

# New Inference: linear combinations of two $\beta_j$

.pull-left-60[
```{r, echo=FALSE}

ggplot(penguins, aes(x = bill_length_mm, 
                     y = body_mass_g, 
                     color = species)) +
theme(legend.position = "bottom")+
geom_point() +
geom_smooth(method = "lm",
             se = FALSE, 
             aes(color=species))+
labs(title = "Regression of mass on bill length and species") 
```
]
.pull-right-40[
.blockquote[
Quick check: does it make sense that the effect of bill length on mass (.bold[slope]) less for Chinstrap than Adelie?]


<br>
.blockquote[
- .bold[Yes:] The green line (chinstrap) is a bit flatter than the red line (adelie)
]

]


---

# New Inference: linear combinations of two $\beta_j$

.blockquote-list[
SE of the Estimated parameter:
$$S E(\hat{\gamma})=\sqrt{c_{i}^{2} \operatorname{Var}\left(\hat{\beta}_{i}\right)+c_{j}^{2} \operatorname{Var}\left(\hat{\beta}_{j}\right)+2 c_{i} c_{j} \operatorname{Cov}\left(\hat{\beta}_{i}, \hat{\beta}_{j}\right)}$$
where $c_{i}$ and $c_{j}$ are known constants.
- "Var" is the variance of an estimator, which equals
$$\operatorname{Var}(\hat{\beta})=S E(\hat{\beta})^{2}$$
- "Cov" is the covariance between two estimators: how do they "co-vary" together?
- an "unscaled" version of correlation
$$\operatorname{Cov}\left(\hat{\beta}_{i}, \hat{\beta}_{j}\right)=\text { how do } \hat{\beta}_{i} \text { and } \hat{\beta}_{j} \text { co-vary? }$$
]

---

# New Inference: linear combinations of two $\beta_j$

.blockquote[
SE of the Estimated parameter:
$$S E(\hat{\gamma})=\sqrt{c_{i}^{2} \operatorname{Var}\left(\hat{\beta}_{i}\right)+c_{j}^{2} \operatorname{Var}\left(\hat{\beta}_{j}\right)+2 c_{i} c_{j} \operatorname{Cov}\left(\hat{\beta}_{i}, \hat{\beta}_{j}\right)}$$
where $c_{i}$ and $c_{j}$ are known constants.
- The effect of bill length on mass for Chinstraps:
$$\gamma=(1) \times \beta_{1}+(1) \times \beta_{4}=\beta_{1}+\beta_{4}$$
where $c_{1}=1$ and $c_{4}=1$.
- The SE of the estimated effect of bill length on mass for Chinstraps:
$$S E(\hat{\gamma})=\sqrt{1^{2} \operatorname{Var}\left(\hat{\beta}_{1}\right)+1^{2} \operatorname{Var}\left(\hat{\beta}_{4}\right)+2(1)(1) \operatorname{Cov}\left(\hat{\beta}_{1}, \hat{\beta}_{4}\right)}$$

]

---

# New Inference: linear combinations of two $\beta_j$

.blockquote-list[
- The SE of the estimated effect of bill length on mass for Chinstraps:

$$SE(\hat{\gamma})=\sqrt{1^{2} \operatorname{Var}\left(\hat{\beta}_{1}\right)+1^{2} \operatorname{Var}\left(\hat{\beta}_{4}\right)+2(1)(1) \operatorname{Cov}\left(\hat{\beta}_{1}, \hat{\beta}_{4}\right)}$$
- Both $\operatorname{Var}\left(\hat{\beta}_{j}\right)$ 's and $\operatorname{Cov}\left(\hat{\beta}_{i}, \hat{\beta}_{j}\right)$ are obtained using the `vcov` command
- `vcov` returns a $(p+1) \times (p+1)$ matrix (rows/cols for each $\hat{\beta}_{j}$ in a model)
- diagonal values are $\operatorname{Var}\left(\hat{\beta}_{j}\right)$'s
- off-diagonal values are $\operatorname{Cov}\left(\hat{\beta}_{i}, \hat{\beta}_{j}\right)$'s
- row/column combination determines "i" and "j"
]

---

class: middle

# New Inference: linear combinations of two $\beta_j$

.blockquote-list.font80[
The SE of the estimated effect of bill length on mass for Chinstraps:

$$SE(\hat{\gamma}) = \sqrt{1^{2} \operatorname{Var}\left(\hat{\beta}_{1}\right)+1^{2} \operatorname{Var}\left(\hat{\beta}_{4}\right)+2(1)(1) \operatorname{Cov}\left(\hat{\beta}_{1}, \hat{\beta}_{4}\right)}\\
\quad \quad = \sqrt{\operatorname{Var}\left(\hat{\beta}_{1}\right)+ \operatorname{Var}\left(\hat{\beta}_{4}\right)+2\operatorname{Cov}\left(\hat{\beta}_{1}, \hat{\beta}_{4}\right)}$$
]

The variance-covariance matrix is:
.code80[
```{r, echo = FALSE}
my.names = c("$\\hat{\\beta}_0$", "$\\hat{\\beta}_1$", "$\\hat{\\beta}_2$", "$\\hat{\\beta}_3$", "$\\hat{\\beta}_4$", "$\\hat{\\beta}_5$")
var.covar <- vcov(peng_interaction_lm) %>% as.data.frame()
r.names <- c("$\\hat{\\beta}_0$", "$\\hat{\\beta}_1$", "$\\hat{\\beta}_2$", "$\\hat{\\beta}_3$", "$\\hat{\\beta}_4$", "$\\hat{\\beta}_5$")
rownames(var.covar) <- my.names
knitr::kable(var.covar, digits = 3, col.names = my.names, row.names = TRUE, escape = FALSE, format = "html")
```
]

---

# New Inference: linear combinations of two $\beta_j$

```{r}
# beta_1 = bill = row/col 2
# beta_4 = bill:chinstrap = row/col 5
vcov(peng_interaction_lm)[c(2,5), c(2,5)]
```

.blockquote.font90[
- $\operatorname{Var}\left(\hat{\beta}_{1}\right)=11.39794^{2}=129.9131$
- $\operatorname{Var}\left(\hat{\beta}_{4}\right)=17.74666^{2}=314.9440$
- $\operatorname{Cov}\left(\hat{\beta}_{1}, \hat{\beta}_{4}\right)=-129.9131$
]


---

# New Inference: linear combinations of two $\beta_j$

The SE of the estimated effect of bill length on mass for Chinstraps:

$$S E(\hat{\gamma})=\sqrt{1^{2} 129.9131+1^{2} 314.9440+2(1)(1)(-129.9131)}=13.60261$$
.font90[
```{r}
sqrt(129.9131 + 314.9440 + 2*(-129.9131))
```
]
.blockquote[
- The estimated value of $\gamma=\beta_{1}+\beta_{4}$ is $59.11774$ with a SE of $13.60261$.
- Is this estimated effect of bill length on mass for Chinstraps statistically significant?
- What is a $95\%$ CI for the true effect?
]

---

# New Inference: linear combinations of two $\beta_j$

.blockquote-list[
- Tests: test stat $t=\frac{\hat{\gamma}-\text { null } \gamma}{S E(\hat{\gamma})}$
- CI: $\hat{\gamma} \pm t * S E(\hat{\gamma})$
- use usual t-distribution with $\mathbf{d f}=n-(p+1)$
- Is the estimated effect of bill length on mass for Chinstraps statistically significant?
$$\begin{array}{l}
H_{0}: \gamma=0 \quad t=\frac{59.11774-0}{13.60261}=4.346 \\
\text {p-value}=2 \times P(T>4.346)=0.00002
\end{array}$$

]

```{r, collapse=TRUE}
(test_stat <- 59.11774/13.60261)
2*pt(-4.346, df = 336)
```

---

# New Inference: linear combinations of two $\beta_j$

.pull-left[

.blockquote[
- Tests: test stat $t=\frac{\hat{\gamma}-\text { null } \gamma}{S E(\hat{\gamma})}$
- CI: $\hat{\gamma} \pm t * S E(\hat{\gamma})$
- use usual t-distribution with df $=n-(p+1)$
- $95 \%$ CI for the effect of bill length on mass for Chinstraps?
$$59.11774 \pm(1.967049)(13.60261)\\=32.36,85.87$$
]


```{r, collapse=TRUE}
qt(.975, df = 336)
```

]
.pull-right[


.blockquote[
- The estimated effect of bill length on mean mass is statistically significant in Chinstraps $(\mathrm{t}=4.346, \mathrm{df}=336, \mathrm{p}<0.0001)$.
- A one $\mathrm{mm}$ increase in bill length in Chinstrap penguins is associated with a $32.4 \mathrm{~g}$ to $85.9 \mathrm{~g}$ increase in mean weight.

<br>
]


```{r, collapse=TRUE}
59.11774 + c(-1,1)*(1.967049)*(13.60261) 

```

]

---

# Mean response and prediction

.blockquote[
- Inference for $\mu_{y \mid x}$ and $\operatorname{pred}(y \mid x)$ are the same in MLR as SLR!!
- Predicting the mass of a Gentoo with a bill length of $40 \mathrm{~mm}$ :
]


```{r}
predict(peng_interaction_lm, 
        newdata = data.frame(bill_length_mm = 40,
                             species = "Gentoo"),
        interval = "prediction")
```


---

# Pengiuns big picture

```{r, echo=FALSE}
library(moderndive)
peng_interaction_lm <-  lm(body_mass_g ~ bill_length_mm*species, data = penguins)
peng_table_interaction <- get_regression_table(peng_interaction_lm)
knitr::kable(peng_table_interaction, digis= 4, format = "html")
```

<br>

.blockquote.out-t.font90[
- What do the last two rows of t-tests?

- What can we tell or not tell from the last two rows of results?
]

---

# Pengiuns big picture


.pull-left[

.out-t.font80[Is the interaction of bill length and species needed?]

```{r, echo=FALSE}
ggplot(penguins, aes(x = bill_length_mm, 
                     y = body_mass_g, 
                     color = species)) +
theme(legend.position = "bottom")+
geom_point() +
geom_smooth(method = "lm",
             se = FALSE, 
             aes(color=species))+
labs(title = "Regression of mass on bill length and species") 
```
]
.pull-right[
.blockquote[

- t-test results for interaction terms tell us *individually* how Chinstrap differs from Adelie $\beta_{4}$ and how Gentoo differs from Adelie $\beta_{5}$
- We can't use t-test results to remove more than one term! 
- All t-test results assume *all other terms are in the model!*

]

]

---

class: middle

# Next...

.blockquote-list[
We will test whether .bold[the effect of bill length on mass depends on species]:

$$H_{0}: \beta_{4}=\beta_{5}=0$$
with a new test done using .bold[ANOVA]!
]


