---
title: "ANOVA for MLR"
subtitle: "<br/> STAT 230"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  xaringan::moon_reader:
    css: ["default", css/xaringan-themer-solns.css, css/my-theme.css, css/my-font.css]
    lib_dir: libs
    chakra: libs/remark-latest.min.js
    seal: false
    nature:
      highlightStyle: googlecode  #http://arm.rbind.io/slides/xaringan.html#77 # idea, magula
      highlightLines: true
      highlightLanguage: ["r", "css", "yaml"]
      countIncrementalSlides: true
      slideNumberFormat: "%current%"
      titleSlideClass: ["left", "middle", "inverse"]
      ratio: "16:9"
    includes:
      in_header: header.html
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
options(htmltools.preserve.raw = FALSE)
options(ggrepel.max.overlaps = Inf)

knitr::opts_chunk$set(echo = TRUE, 
                      dev = 'svg',
                      collapse = FALSE, 
                      comment = NA,  # PRINTS IN FRONT OF OUTPUT, default is '##' which comments out output
                      prompt = FALSE, # IF TRUE adds a > before each code input
                      warning = FALSE, 
                      message = FALSE,
                      fig.height = 3, 
                      fig.width = 4,
                      out.width = "100%"
                      )

# load necessary packages
library(Sleuth3)   # Data-set for Sleuth
library(tidyverse)
library(dplyr)
library(countdown)
library(mosaic)
library(ggthemes)
library(xaringanExtra)
library(forcats)
xaringanExtra::use_panelset()
xaringanExtra::use_tachyons()
xaringanExtra::use_clipboard()
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = TRUE  
)
library(flipbookr)
library(patchwork)
library(DT)
library(moderndive)
library(knitr)
library(grid)
library(gridExtra)
library(Stat2Data)
library(palmerpenguins)

# Set ggplot theme
theme_set(theme_tufte(base_size = 10))

yt <- 0

set.seed(1234)

sim_mlr = function(x1, x2, beta_0 = 2, beta_1 = 3, beta_2 = 2, sigma = 3) {
  n = length(x1)
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x1 + beta_2 * x2 + epsilon
  data.frame(predictor1 = x1, predictor2 = x2, response = y)
}

num_obs = 25
x_vals1 = runif(num_obs, 0, 10)
x_vals2 = rbeta(num_obs, 1,5)*10
sim_data = sim_mlr(x1 = x_vals1, x2 = x_vals2, beta_0 = 3, beta_1 = 2, beta_2= 1.6, sigma = 3)


```


```{r xaringanExtra-clipboard, echo=FALSE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clipboard\"></i>",
    success_text = "<i class=\"fa fa-check\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times-circle\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```


layout: true
  
---

class: title-slide, middle

# .fancy[ANOVA for MLR]

### .fancy[Stat 230]

`r format(Sys.Date(), ' %B %d %Y')`

---

# Overview

.pull-left[
```{r, echo=FALSE, fig.align='center', fig.width=4, fig.height=4, out.width="100%"}
library(gg3D)
ggplot(sim_data, aes(predictor1, y=predictor2, z=response)) + 
  theme_void()+
  axes_3D() +
  stat_3D()
```

]
.pull-right[
<br>
<br>
Today: 
.blockquote-list[
- ANOVA for MLR
- MLR Inference
- $R^2$ adjusted
- ANOVA F-test
]

]

---

# Penguins example

$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}$$


.pull-left[
```{r, echo=FALSE}
ggplot(penguins, aes(x = bill_length_mm, 
                     y = body_mass_g, 
                     color = species)) +
theme(legend.position = "bottom")+
geom_point() +
geom_smooth(method = "lm",
             se = FALSE, 
             aes(color=species))+
labs(title = "Regression of mass on bill length and species") 
```
]
.pull-right[
<br>
.blockquote[
- We will test whether the effect of bill length on mass depends on species:
$$H_{0}: \beta_{4}=\beta_{5}=0$$
]
]

---

# Comparing models

.blockquote[

Is the effect of of bill length the same for Gentoo and Adelie (baseline)?

\begin{align*}
H_0: \beta_{5} &= 0\\
H_A: \beta_{5} &\neq 0
\end{align*}
]

--

<br>

.blockquote.font90[
.bold[Equivalent test:] 
\begin{aligned}
H_{0}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap } \\
H_{A}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}


]

--

.out-t.center[Individual t-tests: Is this term significant assuming all others are in the model?]

---

# Comparing models

.blockquote[
Does the effect of bill length on mass depends on species?

\begin{align*}
H_0: &  \beta_4 = \beta_{5} = 0\\
H_A: & \text{ at least one } \beta_4, \beta_5 \neq 0
\end{align*}
]

--

<br>

.blockquote.font90[
.bold[Equivalent test:]
\begin{aligned}
H_{0}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
H_{A}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}
]

--

.out-t[Our new test ("F-test"): Is at least one of these terms significant assuming all others are in the model?]

---

# Comparing models

\begin{align*}
H_0: & \beta_4 = \beta_{5} = 0\\
H_A: & \text{at least one } \beta_4, \beta_5 \neq 0
\end{align*}





.blockquote.font80[
We can no longer use 

\begin{align*}
t=\frac{\hat{\beta}-0}{S E(\hat{\beta})}
\end{align*}
]

--

<br>

.blockquote-list[

We will use Analysis of Variance (ANOVA)

- .green[idea:] how much more explanatory power does the bigger model $H_{A}$ have compared to the smaller model $H_{0}$ ?
]


---

# ANOVA for MLR (basically same as SLR!)

$$S S T=S S r e g+S S R$$

.blockquote-list.font90[
- SST: Total variation Total sum of squares

$$SST=\sum_{i=1}^{n}\left(y_{i}-\bar{y}\right)^{2}=(n-1) s_{y}^{2}$$

- SSR: Unexplained variation Residual sum of squares

$$S S R=\sum_{i=1}^{n}\left(y_{i}-\hat{y}_{i}\right)^{2}=(n-p-1) \hat{\sigma}^{2}$$
- SSreg: Explained variation Regression sum of squares

$$SSreg=\sum_{i=1}^{n}\left(\hat{y}_{i}-\bar{y}\right)^{2}$$

]

---

# R-squared


```{r, echo=FALSE}
library(palmerpenguins)
library(tidyr) # has `drop_na()` function that drops rows having missing values
penguins <- penguins %>% tidyr::drop_na(bill_length_mm, body_mass_g, species)
peng_bill_lm <-  lm(body_mass_g ~ bill_length_mm, data = penguins)
peng_nointeraction_lm <- lm(body_mass_g ~ bill_length_mm + species, data = penguins)
peng_interaction_lm <-  lm(body_mass_g ~ bill_length_mm*species, data = penguins)

peng_bill_summary <- get_regression_summaries(peng_bill_lm, digits = 5)
peng_bill_table <- get_regression_table(peng_bill_lm, digits = 5)

peng_nointeraction_summary <- get_regression_summaries(peng_nointeraction_lm, digits = 5)
peng_interaction_summary <- get_regression_summaries(peng_interaction_lm, digits = 5)


# bill
ssreg.bill <- sum((peng_bill_lm$fitted.values - mean(peng_bill_lm$fitted.values))^2)
ssr.bill <- sum(((peng_bill_lm$fitted.values - penguins$body_mass_g)^2))
ss.total <- ssreg.bill + ssr.bill

# no interaction
ssreg.nointeraction <- sum((peng_nointeraction_lm$fitted.values - mean(peng_nointeraction_lm$fitted.values))^2)
ssr.nointeraction <- sum(((peng_nointeraction_lm$fitted.values - penguins$body_mass_g)^2))
ss.total <- ssreg.bill + ssr.bill


# interaction

ssreg.interaction <- sum((peng_interaction_lm$fitted.values - mean(peng_interaction_lm$fitted.values))^2)
ssr.interaction <- sum(((peng_interaction_lm$fitted.values - penguins$body_mass_g)^2))
ss.total <- ssreg.bill + ssr.bill

# make data.frame for plotting

my_data <- data.frame(SS = c(ssreg.bill,ssr.bill, ssreg.nointeraction,ssr.nointeraction,ssreg.interaction,ssr.interaction),
            SStype = c("SSreg", "SSR", "SSreg", "SSR", "SSreg", "SSR"),
            model = c("bill \n R^2 = 35.4", "bill \n R^2 = 35.4",
                      "parallel \n R^2 = 78.3", "parallel \n R^2 = 78.3", 
                      "interaction \n R^2 = 78.8", "interaction \n R^2 = 78.8"))


```

.blockquote.font90[
R-squared (coefficient of determination) measures the proportion of variability observed in the response $Y$ which can be explained by the regression of $\mathbf{Y}$ on $x_{1}, \ldots, x_{p}$.
]

<br>

.blockquote-list.font80[

$$R^{2}=\frac{S S r e g}{S S T}=1-\frac{S S R}{S S T}$$

- About $35.4 \%$ of the variation in mass is explained by the regression of mass on bill length.
- About $78.3 \%$ of the variation in mass is explained by the regression of mass on bill length and species.
- About $78.8 \%$ of the variation in mass is explained by the regression of mass on bill length, species and their interaction.

]

---

# Penguins: ANOVA for 3 models

.pull-left-60[

```{r, echo = FALSE}
library(ggthemes)
ggplot(my_data, aes(x = model, y = SS, fill = SStype)) +
  geom_col() +
  labs(y = "Sum of Squares",
       title = "Decomposition of total variation for different models") +
  theme(legend.position = "bottom") +
  scale_fill_wsj() +
  theme(plot.title = element_text(hjust=0.5, size=9, face='bold'))
```


]
.pull-right-40[
<br>

$$R^{2}=\frac{S S r e g}{S S T}=1-\frac{S S R}{S S T}$$

.blockquote.font80[
- SSreg (explained) is largest for the biggest model
- SSR (unexplained) is smallest for the biggest model
]


]

---

# ANOVA for MLR


$$S S T=S S r e g+S S R$$


.blockquote.font90[
- .bold[Mathematically:] bigger models have bigger SSreg:

$$\operatorname{SSreg}\left(x_{1}, x_{2}, x_{1}: x_{2}\right)>\operatorname{SSreg}\left(x_{1}, x_{2}\right)>\operatorname{SSreg}\left(x_{1}\right)$$
- .bold[Mathematically:] bigger models have smaller SSR:

$$S S R\left(x_{1}, x_{2}, x_{1}: x_{2}\right)<\operatorname{SSR}\left(x_{1}, x_{2}\right)<S S R\left(x_{1}\right)$$
- Bigger models have more explanatory power (higher SSreg) and larger $R^{2}$ than a smaller model
- Assume the smaller model is a nested version of the larger
- But is the bigger model "worth it" if it is more complex and only has slightly more SSreg explained?

]

---

# ANOVA for MLR

.out-t[Mean square values divide SS by their degrees of freedom]

.blockquote-list.font80[
- Degrees of freedom (basically) measure how much stuff we've added up in the SS
- .bold[MST:] Average total variation Mean Square of Total
$$M S T=\frac{S S T}{n-1}=s_{y}^{2}=\text { sample variance of } \mathrm{Y}$$
- .bold[MSR:] Average unexplained variation residual mean square
$$M S R=\frac{S S R}{n-(p+1)}=\hat{\sigma}^{2}=\text { estimated model variance }$$
- .bold[MSreg:] Explained variation Regression mean square
$$M S r e g=\frac{S S r e g}{p}$$
]

---

# ANOVA for MLR

$$S S T=S S r e g+S S R$$


.blockquote.font80[
- .bold[Mathematically:] mean square residual and regression values won't equal the total
$$M S T \neq M S r e g+M S R$$
- .bold[Mathematically:] mean square regression values won't be ordered based on their size
$$\operatorname{MSreg}\left(x_{1}, x_{2}, x_{1}: x_{2}\right) ? \operatorname{MSreg}\left(x_{1}, x_{2}\right) ? \operatorname{MSreg}\left(x_{1}\right)$$
- .bold[Mathematically:] mean square residual values won't be ordered based on their size
$$\operatorname{MSR}\left(x_{1}, x_{2}, x_{1}: x_{2}\right) ? \operatorname{MSR}\left(x_{1}, x_{2}\right) ? \operatorname{MSR}\left(x_{1}\right)$$
]

---

class: middle

# Adjusted R-squared

.blockquote.font90[
Adjusted $R^{2}$ is similar to $R^{2}$, but it uses mean square values to account for the size of a model.
$$R_{a}^{2}=R_{\text {adjust }}^{2}=1-\frac{M S R}{M S T}$$

- $R_{a}^{2}$ could actually increase if unneeded predictors are removed from a model.
- $R^{2}$ would never increase when predictors are removed.

]

---

## Penguins: Adjusted $R^2$ for 3 models

.code60[
```{r, collapse=TRUE}
penguins <- penguins %>% tidyr::drop_na(bill_length_mm, body_mass_g, species) # drop missing values
peng_bill_lm <-  lm(body_mass_g ~ bill_length_mm, data = penguins) # simple linear regression model 
peng_nointeraction_lm <- lm(body_mass_g ~ bill_length_mm + species, data = penguins) # parallel lines model
peng_interaction_lm <-  lm(body_mass_g ~ bill_length_mm*species, data = penguins) # interaction model
```
]


.pull-left[

.code60[
```{r, collapse=TRUE}
get_regression_summaries(peng_bill_lm) %>% 
  select(r_squared, adj_r_squared)  # just want these two
```
]


.code60[
```{r, collapse=TRUE}
get_regression_summaries(peng_nointeraction_lm) %>% 
  select(r_squared, adj_r_squared)
```
]

.code60[
```{r, collapse=TRUE}
get_regression_summaries(peng_nointeraction_lm) %>% 
  select(r_squared, adj_r_squared)
```
]
]


.pull-right[

.code90[
```{r, highlight.output = c(3,5,8)}
glimpse(penguins)
```
]
]
---

class: middle

# Penguins: Adjusted $R^2$ for 3 models

.blockquote[
- Removing the interaction terms results in a small drop in adjusted $R^{2}$ (0.785 to $0.781$ ).
- The explanatory power of both the larger interaction model and smaller no-interaction models looks similar.
- .bold[Next:] let's verify this observation with an ANOVA F-test
]

---

# Comparing models: ANOVA F-test

.blockquote.font80[
- .bold[Hypotheses:]
$$\begin{array}{l}
H_{0}: \mu_{Y \mid x}=\text { smaller, nested model } \\
H_{A}: \mu_{Y \mid x}=\text { larger model }
\end{array}$$
where the nested model is arrived at by setting one or more $\beta_{j}$ in the larger model equal to 0
- .bold[Test stat:]
$$F=\frac{\text { extraSS } /(\# \text { terms tested })}{M S R_{\text {larger }}}$$
where extra SS is the difference in explained SS between the larger $H_{A}$ and smaller $H_{0}$ models
$$\begin{aligned}
\text { extraSS } &=S S r e g(\text { larger model })-S S r e g(\text { smaller model }) \\
&=S S R(\text { smaller model })-S S R(\text { larger model })
\end{aligned}$$
]

---

# Comparing models: ANOVA F-test


.pull-left[
```{r, echo=FALSE}
ggplot(data.frame(x = c(0, 5)), aes(x)) +
  stat_function(fun = df,
                geom = "area",
                fill = "turquoise",
                args = list(
                  df1 = 5,
                  df2 = 50
                ))+
  labs(
    title = " F-distribution with DF of 5 and 50",
    x = "F stat values",
    y = "Prob"
  )
```
]

.pull-right[
<br>
.blockquote.font80[
- .bold[P-value:] Probability of getting an T test stat larger than one observed
$$\text{p-value }=P\left(F>f_{\text {observed }}\right)$$
using the F-distribution with degrees of freedom equal to the # of terms tested (numerator) and $n-(p+1)$ from the larger $H_{A}$ model (denominator)
]

]

---

# ANOVA F-test in R


```r
# R-code
anova(smaller_lm, larger_lm)
```



.blockquote[
Let's compare the interaction and parallel lines models in penguins dataset:

$$\begin{aligned}
H_{0}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
H_{A}: \mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
&+\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}$$
- Same as testing $\beta_{4}$ and $\beta_{5}$ in the larger model

$$\begin{array}{l}
H_{0}: \beta_{4}=\beta_{5}=0 \\
H_{A}: \text { at least one } \beta_{4}, \beta_{5} \neq 0
\end{array}$$
]


---


# ANOVA F-test in R

```{r}
anova(peng_nointeraction_lm , peng_interaction_lm)
```


---


# ANOVA F-test in R

```{r, highlight.output = c(5,6,7)}
anova(peng_nointeraction_lm , peng_interaction_lm)
```



.pull-left[
.blockquote[
Res.Df column
- model DF for smaller $=342-4=338$
- model DF for larger $=342-6=336$
]
]
.pull-right[
.blockquote[
RSS Df column
- $\operatorname{RSS}($ smaller $)=47,613,707$
- $\operatorname{RSS}($ larger $)=46,447,006$
]
]

---

# ANOVA F-test in R

```{r, highlight.output = c(5,6,7)}
anova(peng_nointeraction_lm , peng_interaction_lm)
```


.pull-left[
.blockquote[
Df column
- number of terms tested $=6-4=338-336=2$
]
]
.pull-right[
.blockquote[
Sum of Sq column
- Extra SS for this test $=$ RSS(smaller) $-$ RSS(larger $)=1,166,702$
]
]


---

# ANOVA F-test in R

```{r, highlight.output = c(5,6,7)}
anova(peng_nointeraction_lm , peng_interaction_lm)
```


.pull-left[
.blockquote[
F column
- $\mathbf{F}$ test stat $=F=\frac{1,166,702 / 2}{46,447,006 / 336}=4.22$

]
]
.pull-right[
.blockquote[
$\operatorname{Pr}(>F)$ column
- p-value for this test $=P\left(F_{2 ; 336}>4.22\right)=1-\operatorname{pf}(4.22,2,336)\\ = 0.01549$
]
]

---

# Penguins wrapup

```{r, echo=FALSE, out.width="60%", fig.align='center'}
ggplot(penguins, aes(x = bill_length_mm, 
                     y = body_mass_g, 
                     color = species)) +
theme(legend.position = "bottom")+
geom_point() +
geom_smooth(method = "lm",
             se = FALSE, 
             aes(color=species))+
labs(title = "Regression of mass on bill length and species") 
```

.yellow-h.center[The slopes are different enough to warrant an interaction model]

---

# Penguins wrapup


```{r}
summary(peng_interaction_lm)$coefficients # simple way of getting summaries
```


---

# Penguins wrapup

```{r, highlight.output = c(7) }
summary(peng_interaction_lm)$coefficients
```


<br>

.blockquote-list[
- The difference in slopes (effect of bill length on mass) between Adelie (baseline) and Gentoo isn't statistically significance $(\mathrm{t}=0.948, \mathrm{df}$ $=336, \mathrm{p}=0.344)$
]


---


# Penguins wrapup

```{r, highlight.output = c(6) }
summary(peng_interaction_lm)$coefficients
```

<br>

.blockquote-list[

- The difference in slopes (effect of bill length on mass) between Adelie (baseline) and Chinstrap is weakly statistically significance $(\mathrm{t}=-1.994$, $\mathrm{df}=336, \mathrm{p}=0.047$ ).

]


---

# Penguins wrapup

.code80[
```{r, echo=FALSE}
summary(peng_interaction_lm)$coefficients
```
]

$$
\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}+\beta_{1} \text { bill }+\beta_{2} \text { speciesChinstrap }+\beta_{3} \text { speciesGentoo } \\
& \quad +\beta_{4} \text { bill } \times \text { speciesChinstrap }+\beta_{5} \text { bill } \times \text { speciesGentoo }
\end{aligned}
$$

.blockquote-list[
- To compare Chinstrap and Gentoo, either use the test
$$H_{0}: \beta_{1}+\beta_{4}=\beta_{1}+\beta_{5} \Rightarrow \beta_{4}-\beta_{5}=0$$
]


---

# Penguins wrapup

.blockquote[
- OR! Relevel the species variable to make Chinstrap the baseline - it's slope seem most different (flatter) compared to the other two species

$$\begin{aligned}
\mu_{\text {mass } \mid x} &=\beta_{0}^{*}+\beta_{1}^{*} \text { bill }+\beta_{2}^{*} \text { speciesAdelie }+\beta_{3}^{*} \text { speciesGentoo } \\
&+\beta_{4}^{*} \text { bill } \times \text { speciesAlelie }+\beta_{5}^{*} \text { bill } \times \text { speciesGentoo }
\end{aligned}$$
- Same variables but changing up the indicator variables changes our parameter interpretation
- that's why we used $\beta^{*}$ here (not the same values as $\beta$ in the previous model)
- but won't change an inference conclusions or effect estimation!!
]

---

# Penguins wrapup

- OR! Relevel the species variable to make Chinstrap the baseline
- Use `fct_relevel()` from the `forcats` package:

```{r}
library(forcats) 
# just need to specify the first level
penguins$new_species <- fct_relevel(penguins$species, "Chinstrap")
```

Check order using a quick table command:

.pull-left[
```{r, collapse=TRUE}
table(penguins$species)
```
]

.pull-right[
```{r, collapse=TRUE}
table(penguins$new_species) 
```
]

---

# Penguins wrapup

```{r}
peng_interaction_lm_CS <- lm(body_mass_g ~ bill_length_mm * new_species, data = penguins)
summary(peng_interaction_lm_CS)$coefficients
```



---

# Penguins wrapup

```{r, highlight.output = c(6)}
peng_interaction_lm_CS <- lm(body_mass_g ~ bill_length_mm * new_species, data = penguins)
summary(peng_interaction_lm_CS)$coefficients
```



<br>

.blockquote[
- The difference in slopes (effect of bill length on mass) between Chinstrap (baseline) and Adelie is weakly statistically significance $(\mathrm{t}=$ $1.994, \mathrm{df}=336, \mathrm{p}=0.047$ ).
]

---

# Penguins wrapup

```{r, highlight.output = c(7)}
summary(peng_interaction_lm_CS)$coefficients
```

<br>

.blockquote[
- The difference in slopes (effect of bill length on mass) between Chinstrap (baseline) and Gentoo is statistically significance $(\mathrm{t}=2.886, \mathrm{df}$ $=336, \mathrm{p}=0.00416$ ).
]


