---
title: "Interactive Visualizations with Shiny"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      
                      collapse = TRUE, 
                      comment = NA, 
                      prompt = FALSE, 
                      warning = FALSE, 
                      message = FALSE,
                      fig.height = 3, 
                      fig.width = 4,
                      out.width = "100%"
                      )

library(tidyverse)
library(shiny)
library(janitor)
library(purrr)
library(lubridate)
library(plotly)
library(DT)
library(ggthemes)
library(leaflet)
library(maptools)
library(maps)     
library(sp)       
library(maptools) 
library(mapdata)
library(polite)
library(rvest)
library(ggiraph)
library(RColorBrewer)
library(viridis)
```

In this tutorial activity, we will look at some more examples on Shiny implementations.

### 1. Trump tweets: changing variable choice

Consider the following histogram and density curve that shows the distribution of the character counts of Trump tweets:

```{r}
tweets <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/TrumpTweetData.csv")
ggplot(tweets, aes(x = charCount)) + theme_economist_white() +
    geom_histogram(aes(y = ..density..), bins = 50) +
    geom_density(color = "lightgreen", adjust = 2.5)
```

### 1. Histrogram of Trump tweets

Construct the `ui` and `server` objects that render and display the histogram above. Then run `shinyApp(ui, server)` to verify that your app works.

```{r}
ui <- fluidPage(
  radioButtons(inputId = "n_bins", 
               label = "Number of histogram bins", 
               choices = seq(25, 125, by = 25), 
               selected = 50),
  sliderInput(inputId = "bw_adjust", 
             label = "Density smoother bandwidth adjustment", 
             min = 0.1, max = 2, value = 1, step = .1),
  varSelectInput(________, 
              label = _____________, 
              data = select(tweets, charCount, nWords,nRealWords)),
  plotOutput("hist")
)

server <- function(input, output){
  output$hist <- renderPlot({
    ggplot(tweets, aes(x = ______)) +    # the !! (bang bang) unquotes
    geom_histogram(aes(y = ..density..), bins = ______) +
    geom_density(color = "lightgreen", adjust = ______) +
    theme_economist_white() 
  })
}

shinyApp(ui = ui, server = server)
```

\newpage

### 2. Trump tweets: adding a check box to change geometry

```{r}
ui <- fluidPage(
  sliderInput("bw_adjust", label = "Density smoother bandwidth adjustment:",
              min = 0.1, max = 2, value = 1, step = 0.1),
  selectInput("x", label = "Variable:", 
             choices = c("Character count" = "charCount",
                         "Number of words" = "nWords",
                         "Number of real words" = "nRealWords")),
  checkboxInput(______, _______, value=FALSE),
  plotOutput("hist")
)


server <- function(input, output){
  output$hist <-renderPlot({
  if (____________) { 
        ggplot(tweets, _______(y = ______, x="source")) +
          geom_boxplot() + coord_flip()  +
          theme(axis.text=element_text(size=16)) }
  else { 
        ggplot(tweets, _______(x = _______, color="source")) +
          geom_density(adjust = ________, size=1.5) +
          theme(axis.text=element_text(size=16), 
              legend.text=element_text(size=16),
              legend.title=element_text(size=16))  }
})
}

shinyApp(ui = ui, server = server)

```

### 3. Gapminder: scatterplot hovering 

```{r}
gapminder <- read_csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/gapminder2018.csv")

ui <- fluidPage(
    sidebarLayout(
          sidebarPanel(
          # data widget: selecting the variables 
          selectInput("x", label = "x-axis Variable:",
             choices = c("Income" ="income",
                         "Life expectancy" = "life_expectancy",
                         "Population" = "population"), 
             selected="income"),
          selectInput("y", label = "y-axis Variable:",
             choices = c("Income" ="income",
                         "Life expectancy" = "life_expectancy",
                         "Population" = "population"), 
             selected="life_expectancy")
    ),
        mainPanel(
            plotOutput("_______", hover= "hover_coord"), br(),
            DT::dataTableOutput("_______")
        )
    )
)


server <- function(input, output) {
   output$gap_plot <- renderPlot({
   ggplot(gapminder, aes_string(x = ______, y = ______)) + 
    geom_point() + 
    theme(axis.text=element_text(size=16),
          axis.title=element_text(size=16) )
})

   output$hover_data <- renderDataTable({
   nearPoints(gapminder, ___________)[,1:6]
  })
        
}

shinyApp(ui = ui, server = server)

```

Now, repeat the same with `brush_points()`

```{r}
ui <- fluidPage(
    sidebarLayout(
          sidebarPanel(
          # data widget: selecting the variables 
          selectInput("x", label = "x-axis Variable:",
             choices = c("Income" ="income",
                         "Life expectancy" = "life_expectancy",
                         "Population" = "population"), 
             selected="income"),
          selectInput("y", label = "y-axis Variable:",
             choices = c("Income" ="income",
                         "Life expectancy" = "life_expectancy",
                         "Population" = "population"), 
             selected="life_expectancy")
    ),
        mainPanel(
            plotOutput("gap_plot", brush = "plot_brush", height = 400), br(),
            DT::dataTableOutput("table")
        )
    )
)


server <- function(input, output) {
   output$_______ <- renderPlot({
   ggplot(gapminder, aes_string(_________, y = _______)) + 
    geom_point() + 
    theme(axis.text=element_text(size=16),
          axis.title=element_text(size=16) )
})

   output$_______ <- renderDataTable({
   brushedPoints(gapminder, ________)
  })
        
}

shinyApp(ui = ui, server = server)
```


### 4. Interactive graphs with `ggiraph`

```{r}
# Scrape the webpage for tables
table_usafacts <- bow(url = "https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/") %>% scrape() %>% html_elements(css = "table") %>% html_table()

# Extract the right table
covid <- table_usafacts[[2]]

# Clean the data
covid_clean <- covid %>% janitor::clean_names() %>%
  mutate_at(4:6, parse_number) %>% mutate(state = str_to_lower(state))

# Store the geographical coordinates and grouping information
states <- map_data("state")

# Joint the state-wise data for plotting
covid_data <- left_join(______, ______, by = c("region" = "state"))

map1 <- ggplot(covid_data, aes(long, lat, group = group)) + coord_map() + theme_map() +
  geom_polygon_interactive(aes(fill = _____, 
                               tooltip = _____
                               )) + 
  guides(fill=guide_legend(title="______")) + scale_fill_fermenter(type = "div", palette = "Spectral") + theme(legend.position = "right") 
  
ggiraph(code = print(map1))
```


```{r}
covid_data$tooltip <- str_c("State = ", str_to_upper(covid_data$region), "\n Cases =", covid_data$cases)

map2 <- ggplot(covid_data, aes(long, lat, group = group)) + coord_map() + theme_map() +
  geom_polygon_interactive(aes(fill = _______,  tooltip = ______, data_id = _____)) +
  guides(fill=guide_legend(title="______")) + scale_fill_fermenter(type = "div", palette = "Spectral") + 
  theme(legend.position = "right") 
  
ggiraph(code = print(map2), hover_css = "fill:lightgreen;r:10pt;")
```

Now, let's work together to create an interactive Shiny App using `ggiraph`.

```{r}
ui <- fluidPage(
  titlePanel("Tracking Covid in USA"),
  sidebarLayout(
    sidebarPanel(
      radioButtons("variable", "What to project?",
                  choices = _____________,
                  selected = ___________)),
    mainPanel(girafeOutput("plot"))
  )
)

server <- function(input, output) {
    output$plot <- renderGirafe({
        covid_data$tooltip <- str_c("State = ", str_to_upper(covid_data$region), "\n Cases =", covid_data[[input$variable]])

        gg_blah <- ggplot(covid_data, aes(long, lat, group = group)) + coord_map() + theme_map() +
  geom_polygon_interactive(aes(fill = __________,
                               tooltip = _________)) +
  guides(fill=guide_legend(title=str_c(________))) + 
          scale_fill_fermenter(type = "div", palette = "Spectral") + theme(legend.position = "right")
         girafe(ggobj = gg_blah )
      })
}    

app1 <- shinyApp(ui = ui, server = server)
app1
```

### 5. Simple demo with `leaflet`

```{r}

table_usafacts <- bow(url = "https://usafacts.org/visualizations/coronavirus-covid-19-spread-map/state/minnesota") %>% scrape() %>% html_elements(css = "table") %>% html_table()

covidMN <- table_usafacts[[2]]

# tidy it up
covidMN_final <- covidMN %>% janitor::clean_names() %>%
  mutate(cases = as.numeric(str_remove(cases, ","))) %>%
  mutate(county = str_remove(county, " County"))

# leaflet
MNcounty <- map("county","Minnesota", plot=FALSE, fill=TRUE)
MNmap <- map2SpatialPolygons(MNcounty, IDs = MNcounty$names)
map <- SpatialPolygonsDataFrame(MNmap, covidMN_final, match.ID = FALSE)

#pal <- colorNumeric(palette = "viridis", alpha = TRUE, domain = map$cases)

bins <- c(0, 1000, 5000, 10000, 100000, Inf)
pal <- colorBin("viridis", domain = map$cases, bins = bins)

labels <- sprintf("<strong> %s </strong> <br/> Observed: %s", map$county, map$cases) %>%
  lapply(htmltools::HTML)


l <- leaflet(map) %>% addTiles() %>% setView(lng = -93.1616, lat = 44.4583, zoom = 5) 

l %>% addPolygons(color = "grey", weight = 1,
                  fillColor = ~pal(cases), fillOpacity = 0.7,
                  highlightOptions = highlightOptions(weight = 5),
                  label = labels) %>%
                  addLegend(pal = pal, values = ~cases, opacity = 0.5, 
                            title = "Observed Cases", 
                            position = "bottomright")

library(htmlwidgets)
saveWidget(l, file="Minnesota.html")

```

### 6. More with `leaflet`

```{r}
# United States
l <- leaflet() %>% addTiles() %>% setView(-96, 37.8, 4) 

reg <- "(?=:).+"
str_remove("sdf:df", reg)

table_usafacts <- bow(url = "https://usafacts.org/visualizations/covid-vaccine-tracker-states/state/minnesota") %>%
  scrape() %>% html_elements(css = "table") %>% html_table()

covid <- table_usafacts[[3]]

# Clean the data

covid_final <- covid %>% janitor::clean_names() %>%
  mutate_at(2:4, parse_number) %>% mutate(state = str_to_lower(state))%>%
  filter(state %in% c("minnesota", "wisconsin", "iowa", "michigan", "illinois", "indiana"))

USA <- maps::map("state", regions = c("minnesota", "wisconsin", "iowa", "michigan", "illinois","indiana"), plot = FALSE, fill=TRUE)

USA$names <- str_remove(USA$names, reg)
Regionmap <- map2SpatialPolygons(USA, IDs = USA$names)
map <- SpatialPolygonsDataFrame(Regionmap, covid_final, match.ID = FALSE)

bins <- seq(min(map$percent_fully_vaccinated), max(map$percent_fully_vaccinated), length.out = 6)
pal <- colorBin("viridis", domain = map$percent_fully_vaccinated, bins = bins)

labels <- sprintf("<strong> %s </strong> <br/> Observed: %s", str_to_upper(map$state), map$percent_fully_vaccinated) %>%
  lapply(htmltools::HTML)


m <- leaflet(map) %>% addTiles() %>% setView(lng = -93.1616, lat = 44.4583, zoom = 4)

n <- m %>% addPolygons(color = "grey", weight = 1,
                  fillColor = ~pal(percent_fully_vaccinated), fillOpacity = 0.7,
                  highlightOptions = highlightOptions(weight = 5),
                  label = labels) %>%
                  addLegend(pal = pal, values = ~percent_fully_vaccinated, opacity = 0.5, 
                            title = "Percent Vaccn.", 
                            position = "bottomright")


library(htmlwidgets)
saveWidget(n, file="USA.html")
```

