---
title: "Graphics with ggplot2"
subtitle: "<br/> STAT 220"
author: "Bastola"
date: "`r format(Sys.Date(), ' %B %d %Y')`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      ratio: '16:9'
      titleSlideClass: ["center", "top", "my-title"]
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---

<style type="text/css">

.hljs-github .hljs {
    background: #e5e5e5;
}

.inline-c, remark-inline-code {
   background: #e5e5e5;
   border-radius: 3px;
   padding: 4px;
   font-family: 'Source Code Pro', 'Lucida Console', Monaco, monospace;
}

.yellow-h{
   background: #ffff88;
}

.out-t, remark-inline-code {
   background: #9fff9f;
   border-radius: 3px;
   padding: 4px;
   
}

.pull-left-c {
    float: left;
    width: 58%;
    
}

.pull-right-c {
    float: right;
    width: 38%;
    
}

.medium {
    font-size: 75%
    
}

.small {
    font-size: 50%
    }

.action {
    background-color: #f2eecb;
  
}

</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      dev = 'svg',
                      collapse = TRUE, 
                      comment=NA, 
                      warning = FALSE,
                      message = FALSE,
                      fig.height = 4, fig.width = 6, fig.align='center')



library(dplyr)
library(ggplot2)
library(countdown)
library(ggthemes)
library(tidyverse)
library(dslabs)
library(datasauRus)
```

```{css, echo = FALSE}
.red{ color: red;}
.remark-code, .remark-inline-code {
    font-family: 'Fira Mono', 'Source Code Pro', monospace;}
.remark-slide-scaler {
    overflow-y: auto;
}
```


# Layered Grammar of Graphics

.pull-left[
- Essentials

    - Data

    - **Aesthetic mappings**

    - **Geometric objects**

]

--

.pull-right[

- Additional elements

    - Facets

    - **Coordinate system**

    - **Statistical transformations**

    - **Position adjustments**

    - Scales

    - Theme

]

---

# Common **ggplot2** options 


```
ggplot(data) +    # data
  <geometry_funs>(aes(<variables>)) + # aesthetic variable mapping
  <label_funs> +  # add context
  <facet_funs> +  # add facets (optional)
  <coordinate_funs> +  # play with coords (optional)
  <scale_funs> + # play with scales (optional)
  <theme_funs> # play with axes, colors, etc (optional)
```

--

- See the [Rstudio cheatsheets](https://rstudio.com/resources/cheatsheets/) for more details


---

# Hate Crime and income inequality

A FiveThirtyEight article published in 2017 claimed that higher rates of hate crimes were tied to greater income inequality.

```{r,  warning=FALSE, message=FALSE, echo=-1}
hate_crimes <- read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/HateCrime.csv")
glimpse(hate_crimes)
```

---

# Layering geoms


When you are iteratively building plots, it's useful to store the .yellow-h[base plot] as an object


```{r fig.height = 3.7, fig.wigth = 7, fig.align='center', eval=FALSE}
base <- ggplot(hate_crimes, aes(x=median_household_income, y=hate_crimes_per_100k_splc)) 
base
```

--

```{r fig.height = 3.7, fig.wigth = 7, fig.align='center', echo=FALSE}
base <- ggplot(hate_crimes, aes(x=median_household_income, y=hate_crimes_per_100k_splc)) 
base
```

---

# Layering geoms

```{r fig.height = 3.7, fig.wigth = 7, fig.align='center', eval=FALSE}
base + 
  geom_point() 
```

--

```{r fig.height = 3.7, fig.wigth = 7, fig.align='center', echo=FALSE}
base + 
  geom_point() 
```

---

# A better plot 

```{r, warning=FALSE}
base + 
  geom_text(aes(label=state_code, color=region))
```

---

# Labeling your graphics

In `ggplot2` you can add/change the title, subtitle, caption, and x- and y-axis labels by adding a `labs()` layer. Below is an example illustrating it’s use:

```{r eval=FALSE}
ggplot(data = mpg) +
geom_point(mapping = aes(x = displ, y = hwy)) + labs(
    title = "Put your informative title here",
    subtitle = "and your subtitle here",
    x = "New x label",
    y = "New y label",
    caption = "Put a caption here"
  )
```

---


# Plotting variables

```{r fig.height = 4, fig.width = 6, fig.align='center'}
ggplot(data = mtcars, aes(x=mpg, y=cyl)) + 
  geom_point(aes(col=as.factor(cyl), pch=as.factor(cyl))) +
  labs(x='Miles per gallon', y='Number of cylinders') + 
  guides(col=FALSE, pch=FALSE)
```


---
class: action

# Your Turn 1
Please git clone the GitHub repository [`03-visualizations`](https://github.com/stat220/03-visualization-activity).

The data frame in this exercise is called `datasaurus_dozen` and it's in the `datasauRus` package. This single data frame contains 13 datasets. 

.pull-left[

```{r, echo=FALSE}
dino_data <- datasaurus_dozen %>%
   filter(dataset == "dino")
ggplot(data = dino_data, mapping = aes(x = x, y = y)) +
  geom_point()
```

]

.pull-right[
`r countdown(minutes = 5, seconds = 00, top = 1, color_background = "inherit")`
]

---

# Statistical transformations: Default stats

Common geom | stat 
---------------|-----------------
`geom_histogram()`|`stat_bin()`     
`geom_bar()`|`stat_count()`     
`geom_smooth()`|`stat_smooth()`  
`geom_boxplot()`|`stat_boxplot()` 
`geom_density()`|`stat_density()`   

--

- Every geom has a default stat

- Often no need to explicitly specify the stat

- Check help files: `?geom_bar`

---

# Weather Dataset

The data set `Weather` contains data on weather-related variables for several world cities.

```{r}
#install.packages(mosaicData)
library(mosaicData)
data(Weather)
glimpse(Weather)
```

---

# Weather patterns

```{r}
ggplot(Weather, aes(y=avg_temp)) + geom_boxplot()
```

---

# Weather patterns

```{r}
ggplot(Weather, aes(y=avg_temp, group=city)) + geom_boxplot(aes(color=city))
```

---

# Weather patterns

```{r, fig.height = 3.7, fig.wigth = 7, fig.align='center'}
ggplot(Weather, aes(x=date, y=avg_temp))+geom_point()
```

---

# Weather patterns

```{r, fig.height = 3.7, fig.wigth = 7, fig.align='center'}
ggplot(Weather, aes(x=date, y=avg_temp)) + 
  geom_point(aes(color=city))
```

---

# Weather patterns in Chicago

```{r, fig.height = 3.7, fig.wigth = 7, fig.align='center'}
Chicago <- Weather %>% filter(city=='Chicago')
ggplot(Chicago, aes(x=date, y=avg_temp)) + 
  geom_point(color='blue')
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=date, y=avg_temp)) + 
  geom_point(aes(color=events))
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=date, y=avg_temp)) + 
  geom_smooth()
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=date, y=avg_temp)) + 
  geom_smooth()+ geom_point()
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=low_temp, y=high_temp)) + 
  geom_point(aes(color=month))
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=low_temp, y=high_temp)) + 
  geom_point(aes(color=as.factor(month)))
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp)) + 
  geom_histogram() 
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp)) + 
  geom_histogram(bins = 50) 
```

---





# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp)) + 
  geom_density(fill='lightblue')
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp)) + 
  geom_density(fill='lightblue', alpha=0.5)
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp, group=month)) +
  geom_density(aes(color=as.factor(month), fill=as.factor(month)), alpha=0.5)
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_temp, group=month)) +
  geom_density(aes(color=as.factor(month), fill=as.factor(month)), alpha=0.5) +
  facet_wrap(~month, nrow=3)
```

---

# Weather patterns in Chicago

```{r}
ggplot(Chicago, aes(x=avg_vis, group=month)) +
  geom_density(aes(color=as.factor(month), fill=as.factor(month)), alpha=0.5) +
  facet_wrap(~month, nrow=3)
```

---
class: action

# Your Turn 2

- `datasaurus_dozen` in the `datasauRus` package contains 13 datasets. We will use 4 of these datasets for this exercise.

-  Create this ribbon of plots using `facet_wrap`.

```{r, echo=FALSE}
library(datasauRus)
dino_data_four <- datasaurus_dozen %>%
   filter(dataset == "dino" | dataset == "star" | dataset == "x_shape"| dataset == "away")
```

```{r, echo=FALSE}
ggplot(dino_data_four, aes(x = x, y = y, color = dataset)) +
  geom_point() +
  facet_wrap(~ dataset, ncol = 2) +
  theme(legend.position = "none")
```

`r countdown(minutes = 5, seconds = 00, left = 0 , color_background = "inherit")`

---

# Weather events

```{r}
ggplot(Weather, aes(x=events)) + 
  geom_bar()+
  theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust=1))
```

---

# Weather events &mdash; proportions

```{r}
ggplot(Weather, aes(x=events, y = ..count../sum(..count..))) +   # change y-axis to proportion
  geom_bar()+
  theme(axis.text.x = element_text(angle = 15, vjust = 0.5, hjust=1))
```

---

# Weather events

```{r}
ggplot(Weather, aes(x=events)) + 
  geom_bar() + 
  coord_flip()
```

---

# Weather events in Chicago

```{r}
ggplot(Chicago, aes(x=events)) + 
  geom_bar() + 
  coord_flip() +
  facet_wrap(~year, nrow=2)
```


---
class: action

# Your Turn 3

- The `mpg` data set is loaded with the `tidyverse`. Run `?mpg` for info.

-  Create this bar chart of vehicle `class`

```{r echo=FALSE}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x = class, fill = class))
```

`r countdown(minutes = 3, seconds = 00, top = 0, color_background = "inherit")`

---

# Gapminder dataset

```{r}
gap_dat <-read.csv("https://raw.githubusercontent.com/deepbas/statdatasets/main/gapminder2018.csv")
gapminder <- gap_dat %>% filter(year == 2018)
glimpse(gapminder)

```


---
class: action

# Your Turn 4

The provides values for life expectancy, GDP per capita, and population, every five years, from 1960 to 2018.
Use `gapminder` dataset to answer the given set of problems.

```{r, echo=FALSE}
gap_dat <- read_csv("/home/deepak/Desktop/stat220/data/gapminder2018.csv")
gapminder <- gap_dat %>% filter(year == 2018)

ggplot(gapminder, mapping = aes(x = four_regions, y = population)) +
  geom_violin() +
  scale_y_log10()
```


`r countdown(minutes = 10, seconds = 00, left = 0, color_background = "inherit")`

