---
title: "Simple R Markdown Example"
author: "Deepak Bastola"
date: 2020-12-01T21:13:14-05:00
categories: ["R"]
tags: ["R Markdown"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# R Markdown

This is a R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

You can use asterisk mark to provide emphasis, such as `*italics* or **bold**`.

You can create lists with a dash:

```
- Item 1
- Item 2
- Item 3
  + Subitem 1
* Item 4
```

- Item 1
- Item 2
- Item 3
  + Subitem 1
* Item 4


You can embed Latex equations in-line, $\frac{1}{n} \sum_{i=1}^{n} x_{i}$ or in a new line as

$$\text{Var}(X) = \frac{1}{n-1}\sum_{i-1}^{n} (x_{i} - \bar{x})^2 $$
## Embed an R code chunk:

```
Use back ticks to 
create a block of code
```

You can also evaluate and display the results of R code. Each tasks can be accomplished in a suitably labeled chunk like the following:

```{r cars}
summary(cars)
fit <- lm(dist ~ speed, data = cars)
fit
```


## Including Plots

You can also embed plots. See Figure \@ref(fig:pie) for example:

```{r pie, fig.cap='A fancy pie chart.', fig.height=7, fig.width=7}
par(mar = c(0, 1, 0, 1))
pie(
  c(280, 60, 20),
  c('Sky', 'Sunny side of pyramid', 'Shady side of pyramid'),
  col = c('#0292D8', '#F7EA39', '#C4B632'),
  init.angle = -50, border = NA
)
```

## Read in data files

```{r read_data}
simple_data <- read.csv("https://deepbas.io/data/simple-1.dat", )
glimpse(simple_data) 
```


```{r pretty_table}
knitr::kable(simple_data, format = "html")
```

## Hide the code

If we enter the `echo = FALSE` option in the R chunk (see the .Rmd file). This prevents the R code from being printed to your document; you just see the results.

```

```{r pretty_table, echo=FALSE}
knitr::kable(simple_data, format = "html")
```
```

```{r pretty_table}
knitr::kable(simple_data, format = "html")
```

